### 2016-03-01 Manuscript/SW_early_life_telomeres_tables_and_figures.Rmd
More edits (RTL)

**Figure 2** Temporal variation in early-life telomere length in the Seychelles warbler based on cross-sectional data. Points and error bars represent mean and standard error of telomere length for all birds born in each summer breeding season. Sample sizes for each year are provided in brackets.

**Figure 3** Factors affecting early-life telomere length in Seychelles warblers at the individual level, within seasons. Model averaged estimates and 95% confidence intervals for explanatory terms used in a linear mixed model with telomere length (cross-sectional data, **A**) and telomere loss (longitudinal data, **B**) as the response variable. Numbers in brackets are the relative importance of each term in the top model set (see main text for details; TQ = territory quality, TL = telomere length).

**Figure 4** Telomere length in relation to **A** season and **B** tarsus length in juvenile Seychelles warblers. In *A**, white and grey boxes represent summer and winter seasons, respectively, while in **B** black and grey points represent nestlings and fledglings, respectively. Sample sizes for each group are provided in brackets.

**Figure 5** Kaplan-Meier curves showing the relationship between survival and telomere length for juvenile Seychelles warblers. Telomere length is binned into groups here for visualisation purposes only (long and short = greater than or less than median telomere length, respectively).

@@

Fig1A <- ggplot(temp,aes(x = Agemonths,y = LogTL))+
  geom_jitter(col='grey') +
  ylab('Log telomere length (kb)') + 
  xlab('Age (months)') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        legend.position = 'none')+
  geom_boxplot(outlier.colour=NA)+
  annotate('text',x = 0.6,y = 4.8,label = 'A')

#
temp <- Loss
temp[,'Age (months)'] <- factor(temp$Agemonths)

Fig1B <- ggplot(temp,aes(x = LogTL,y = LogTL1,col = `Age (months)`,shape = `Age (months)`)) +
  ylab('Log adult telomere length (kb)') + 
  xlab('Log juvenile telomere length (kb)') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        legend.position = 'none')+
  geom_point() +
  stat_smooth(method = 'lm')+
  xlim(2,4.5)+
  ylim(2,4.5)+
    scale_colour_manual(values = c('darkgrey','darkgreen','darkorange'))+
  annotate('text',x = 2,y = 4.5,label = 'B')

#
temp <- Loss
temp[,'Age (months)'] <- factor(temp$Agemonths)

Fig1C <- ggplot(temp,aes(x = TimeDiff,y = Loss,col=`Age (months)`,shape = `Age (months)`)) +
  ylab('Log telomere loss (kb)') + 
  xlab('Time interval (years)') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        legend.position = 'none')+
  geom_point()+
  geom_abline(slope = 0,intercept = 0,lty = 2)+
  stat_smooth(method='lm')+
  scale_colour_manual(values = c('darkgrey','darkgreen','darkorange'))+
  annotate('text',x = 0,y = 2,label = 'C')

Fig1D <- ggplot(temp,aes(x = TimeDiff,y = TROC,col=`Age (months)`,shape = `Age (months)`)) +
  ylab('Log telomere loss rate (kb/year)') + 
  xlab('Time interval (years)') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        legend.position = c(0.9,0.8))+
  geom_point()+
  geom_abline(slope = 0,intercept = 0,lty = 2)+
  scale_colour_manual(values = c('darkgrey','darkgreen','darkorange'))+
  annotate('text',x = 0,y = 8,label = 'D')

pdf('Figures/Figure 1.pdf',
    width = 8,
    height = 8,
    paper = 'a4r')
multiplot(Fig1A,Fig1B,Fig1C,Fig1D,layout=matrix(c(1:4),nrow = 2,byrow=T))
dev.off()

@@

###############################
temp <- juv
temp$LayYear <- factor(temp$LayYear)

Fig2A <- ggplot(subset(temp,Agemonths == 1),
       aes(x = LayYear,
           y = LogTL)) +
  ylab('') + 
  xlab('') +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 60,size = 14,vjust=0.5), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = 'none') +
  geom_jitter(col = 'grey')+
  ylim(c(2,4.5))+
  geom_boxplot(outlier.colour = NA)+
  scale_x_discrete(limits = factor(1998:2014),drop=F)+
  annotate('text',x = 1,y = 4.4,label = 'A')

Fig2B <- ggplot(subset(juv,Agemonths == 6),
       aes(x = factor(LayYear),
           y = LogTL)) +
  ylab('Log telomere length (kb)') + 
  xlab('') +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 60,size = 14,vjust=0.5), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = 'none') +
  geom_jitter(col = 'grey')+
  ylim(c(2,4.5))+
  geom_boxplot(outlier.colour = NA)+
  scale_x_discrete(limits = factor(1998:2014),drop=F)+
  annotate('text',x = 1,y = 4.4,label = 'B')

Fig2C <- ggplot(subset(juv,Agemonths == 10),
       aes(x = factor(LayYear),
           y = LogTL)) +
  ylab('') + 
  xlab('Cohort') +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 60,size = 14,vjust=0.5), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = 'none') +
  geom_jitter(col = 'grey')+
  ylim(c(2,4.5))+
  geom_boxplot(outlier.colour = NA)+
  scale_x_discrete(limits = factor(1998:2014),drop=F)+
  annotate('text',x = 1,y = 4.4,label = 'C')

pdf('Figures/Figure 2.pdf',
    width = 6,
    height = 8,
    paper = 'a4r')
multiplot(Fig2A,Fig2B,Fig2C)
dev.off()

@@

@@

#Run the model
lm2 <- lmer(LogTL ~ Agemonths * (Tarsus + Helper + NonHelper + Sex + BodyMass  + TQ)  + (1|LayYear) + (1|PlateID) + (1|TerritoryID),
            data = temp,
            REML = FALSE)
lm2 <- standardize(lm2)
#Model averaging

@@

#Sort out row/col names
nams <- rownames(table2)
nams[nams == 'z.NonHelper'] <- 'Non helpers'
nams[nams == 'c.Sex'] <- 'Sex'
nams[nams == 'z.TQ'] <- 'TQ'
nams[nams == 'c.Helper'] <- 'Helpers'
nams[nams == 'z.BodyMass'] <- 'Body mass'
nams[nams == 'z.Agemonths'] <- 'Age'
nams[nams == 'z.Tarsus'] <- 'Tarsus'
nams[nams == 'z.Agemonths:z.NonHelper'] <- 'Age x non helpers'
nams[nams == 'c.Sex:z.Agemonths'] <- 'Age x sex'
nams[nams == 'z.Agemonths:z.TQ'] <- 'Age x TQ'
nams[nams == 'c.Helper:z.Agemonths'] <- 'Age x helpers'
nams[nams == 'z.Agemonths:z.BodyMass'] <- 'Age x mass'
nams[nams == 'z.Agemonths:z.Tarsus'] <- 'Age x tarsus'

@@

Fig3 <- forestplot(table2)+
  ylim(-0.4,0.4)

pdf('Figures/Figure 3.pdf',
    width = 8,
    height = 5,
    paper = 'a4r')
Fig3
dev.off()

@@

```{r Figure 4}
temp <- juv

Fig4A <- ggplot(temp,
       aes(x = factor(Agemonths),
           y = LogTL,
           fill=`Non helpers`)) +
  ylab('Log telomere length (kb)') + 
  xlab('Age (months)') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        legend.position = 'top') +
  geom_jitter(aes(colour = `Non helpers`),position=position_jitterdodge(dodge.width = 0.8))+
  geom_boxplot(outlier.colour = NA,notch=F)+
  scale_colour_manual(values = c('grey','darkorange'))+
  scale_fill_manual(values = c('grey','darkorange'))+
  annotate('text',x = 0.55,y = 4.5,label = 'A')

Fig4B <- ggplot(temp,
       aes(x = factor(Agemonths),
           y = LogTL,
           fill=`Helpers`)) +
  ylab('Log telomere length (kb)') + 
  xlab('Age (months)') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        legend.position = 'top') +
  geom_jitter(aes(colour = `Helpers`),position=position_jitterdodge(dodge.width = 0.8))+
  geom_boxplot(outlier.colour = NA,notch=F)+
  scale_colour_manual(values = c('grey','darkorange'))+
  scale_fill_manual(values = c('grey','darkorange'))+
  annotate('text',x = 0.55,y = 4.5,label = 'B')

Fig4C <- ggplot(temp,
       aes(x = TQ,
           y = LogTL,
           colour=`Age (months)`,
           fill = `Age (months)`,
           shape = `Age (months)`)) +
  ylab('Log telomere length (kb)') + 
  xlab('Territory quality') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        legend.position = 'top') +
  geom_point(alpha = 0.2)+
  stat_smooth(method = 'lm')+
  scale_colour_manual(values = c(grey(0.2),'darkorange','darkgreen'))+
  scale_fill_manual(values = c('darkgrey','darkorange','darkgreen'))+
  annotate('text',x = 3,y = 4.5,label = 'C')

Fig4D <- ggplot(temp,
       aes(x = Tarsus,
           y = LogTL,
           colour=`Age (months)`,
           fill = `Age (months)`,
           shape = `Age (months)`)) +
  ylab('Log telomere length (kb)') + 
  xlab('Tarsus length (mm)') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        legend.position = 'top') +
  geom_point(alpha = 0.2)+
  stat_smooth(method = 'lm')+
  scale_colour_manual(values = c(grey(0.2),'darkorange','darkgreen'))+
  scale_fill_manual(values = c('darkgrey','darkorange','darkgreen'))+
  annotate('text',x = 17,y = 4.5,label = 'D')

pdf('Figures/Figure 4.pdf',
    width = 7,
    height = 8,
    paper = 'a4r')
multiplot(Fig4A,Fig4B,Fig4C,Fig4D,layout=matrix(c(1:4),byrow = TRUE,nrow = 2))
dev.off()

```{r Figure 5}
temp <- subset(juv,LayYear<2009)
mymed <- median(temp$cohortTL)
temp$coTLF <- ifelse(temp$cohortTL>mymed,'h','l')

s1 <- survfit(Surv(time = Lifespan, event = Died)~coTLF, data = temp)
Fig5A <- ggsurv(s1,plot.cens=F,lty.est=c(1:2)) + 
  theme_classic() +
  ylab('Proportion survived') + 
  xlab('Age (years)') + 
  xlim(c(0,15)) +
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = 'none') +
  scale_colour_manual(name = 'TL',
                      values = c('grey','black')) +
  guides(colour=F)

mymed <- median(temp$cenTL)
temp$cenTLF <- ifelse(temp$cenTL>mymed,'h','l')
s1 <- survfit(Surv(time = Lifespan, event = Died)~cenTLF, data = temp)
Fig5B <- ggsurv(s1,plot.cens=F,lty.est=c(1,2)) + 
  theme_classic() +
  ylab('Proportion survived') + 
  xlab('Age (years)') + 
  xlim(c(0,15)) +
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = 'none') +
  scale_colour_manual(name = 'TL',
                      values = c('black','grey')) +
  guides(colour=F)

pdf('Figures/Figure 5.pdf',
    width = 4,
    height = 8,
    paper = 'a4')
multiplot(Fig5A,Fig5B)
dev.off()

>>>>

**Figure 2** Factors affecting early-life telomere length in Seychelles warblers. **A** Model averaged estimates and 95% confidence intervals for explanatory terms used in a linear mixed model with telomere length (see main text for details; TQ = territory quality, TL = telomere length).

**Figure 3** Factors affecting early-life telomere shortening rates in Seychelles warblers, within seasons. Model averaged estimates and 95% confidence intervals for explanatory terms used in a linear mixed model with telomere length (cross-sectional data, **A**) and telomere loss (longitudinal data, **B**) as the response variable. Numbers in brackets are the relative importance of each term in the top model set (see main text for details; TQ = territory quality, TL = telomere length).

**Figure 4** Kaplan-Meier curves showing the relationship between survival and telomere length for juvenile Seychelles warblers. Telomere length is binned into groups here for visualisation purposes only (long and short = greater than or less than median telomere length, respectively).

@@

Fig1A <- ggplot(temp,aes(x = Agemonths,y = RTL))+
  geom_jitter(col='grey') +
  ylab('RTL') + 
  xlab('Age (months)') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        legend.position = 'none')+
  geom_boxplot(outlier.colour=NA,alpha = 0.2)+
  annotate('text',x = 0.6,y = 2.1,label = 'A')

temp <- Loss
temp$Agemonths <- factor(temp$Agemonths)

Fig1B <- ggplot(temp,aes(x = Agemonths,y = TROC))+
  geom_jitter(col='grey') +
  ylab('Loss in RTL') + 
  xlab('Age (months)') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        legend.position = 'none')+
  geom_boxplot(outlier.colour=NA,alpha = 0.2)+
  annotate('text',x = 0.6,y = 2.1,label = 'B')

temp <- juv
temp$LayYear <- factor(temp$LayYear)

Fig1C <- ggplot(temp,
       aes(x = LayYear,
           y = RTL)) +
  ylab('RTL') + 
  xlab('') +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 60,size = 14,vjust=0.5), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = 'none') +
  geom_jitter(col = 'grey')+
  geom_boxplot(outlier.colour = NA,alpha = 0.2)+
  scale_x_discrete(limits = factor(1998:2014),drop=F)+
  annotate('text',x = 1,y = 2.1,label = 'C')

pdf('Figures/Figure 1.pdf',
    width = 6,
    height = 6,
    paper = 'a4r')
multiplot(Fig1A,Fig1B,Fig1C,layout=matrix(c(1,2,3,3),nrow = 2,byrow=T))
dev.off()

@@

#####################
#2A
#####################

#Run the model
lm2 <- lmer(RTL ~ Agemonths * (Tarsus + Helper + NonHelper + Sex + BodyMass  + TQ + Season)  + (1|LayYear) + (1|TerritoryID),
            data = temp,
            REML = FALSE)
lm2 <- standardize(lm2)
#Model averaging

#Sort out row/col names
nams <- rownames(table2)
nams[nams == 'z.NonHelper'] <- 'Non helpers'
nams[nams == 'c.Sex'] <- 'Sex'
nams[nams == 'z.TQ'] <- 'TQ'
nams[nams == 'c.Helper'] <- 'Helpers'
nams[nams == 'z.BodyMass'] <- 'Body mass'
nams[nams == 'z.Agemonths'] <- 'Age'
nams[nams == 'z.Tarsus'] <- 'Tarsus'
nams[nams == 'c.Season'] <- 'Season'
nams[nams == 'c.Season:z.Agemonths'] <- 'Age x season'
nams[nams == 'z.Agemonths:z.NonHelper'] <- 'Age x non helpers'
nams[nams == 'c.Sex:z.Agemonths'] <- 'Age x sex'
nams[nams == 'z.Agemonths:z.TQ'] <- 'Age x TQ'
nams[nams == 'c.Helper:z.Agemonths'] <- 'Age x helpers'
nams[nams == 'z.Agemonths:z.BodyMass'] <- 'Age x mass'
nams[nams == 'z.Agemonths:z.Tarsus'] <- 'Age x tarsus'

Fig2A <- forestplot(table2)+
  annotate('text',x = 1,y = 0.25,label = 'A')

Fig2B <- ggplot(temp,
       aes(x = factor(Agemonths),
           y = RTL,
           fill=Season)) +
  ylab('RTL') + 
  xlab('Age (months)') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        legend.position = 'top') +
  geom_jitter(aes(colour = Season),position=position_jitterdodge(dodge.width = 0.8))+
  geom_boxplot(outlier.colour = NA,notch=F,alpha=0.2)+
  scale_colour_manual(values = c('grey','darkorange'))+
  scale_fill_manual(values = c('grey','darkorange'))+
  annotate('text',x = 0.55,y = 2.1,label = 'B')

Fig2C <- ggplot(temp,
       aes(x = Tarsus,
           y = RTL,
           colour=`Age (months)`,
           fill = `Age (months)`,
           shape = `Age (months)`)) +
  ylab('RTL') + 
  xlab('Tarsus length (mm)') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        legend.position = 'top') +
  geom_point(alpha = 0.2)+
  stat_smooth(method = 'lm')+
  scale_colour_manual(values = c(grey(0.2),'darkorange','darkgreen'))+
  scale_fill_manual(values = c('darkgrey','darkorange','darkgreen'))+
  annotate('text',x = 17,y = 2.1,label = 'C')

pdf('Figures/Figure 2.pdf',
    width = 8,
    height = 9,
    paper = 'a4r')
multiplot(Fig2A,Fig2B,Fig2C,layout = matrix(c(1,1,2,3),nrow = 2,byrow = T))
dev.off()

@@

temp <- Loss

@@

#Run the model
lm2 <- lmer(TROC ~ Agemonths * (Tarsus + Helper + NonHelper + Sex + BodyMass  + TQ + Season) + (1|LayYear),
            data = temp,
            REML = FALSE)
lm2 <- standardize(lm2)
#Model averaging

@@

#Sort out row/col names
nams <- rownames(table2)
nams[nams == 'z.NonHelper'] <- 'Non helpers'
nams[nams == 'c.Sex'] <- 'Sex'
nams[nams == 'z.TQ'] <- 'TQ'
nams[nams == 'c.Helper'] <- 'Helpers'
nams[nams == 'z.BodyMass'] <- 'Body mass'
nams[nams == 'z.Agemonths'] <- 'Age'
nams[nams == 'z.Tarsus'] <- 'Tarsus'
nams[nams == 'c.Season'] <- 'Season'
nams[nams == 'c.Season:z.Agemonths'] <- 'Age x season'
nams[nams == 'z.Agemonths:z.NonHelper'] <- 'Age x non helpers'
nams[nams == 'c.Sex:z.Agemonths'] <- 'Age x sex'
nams[nams == 'z.Agemonths:z.TQ'] <- 'Age x TQ'
nams[nams == 'c.Helper:z.Agemonths'] <- 'Age x helpers'
nams[nams == 'z.Agemonths:z.BodyMass'] <- 'Age x mass'
nams[nams == 'z.Agemonths:z.Tarsus'] <- 'Age x tarsus'

@@

Fig3A <- forestplot(table2)+
  annotate('text',x = 1,y = 0.45,label = 'A')

temp <- Loss

Fig3B <- ggplot(temp,
       aes(x = factor(Agemonths),
           y = TROC,
           fill=Season)) +
  ylab('Loss in RTL') + 
  xlab('Age (months)') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        legend.position = 'top') +
  geom_jitter(aes(colour = Season),position=position_jitterdodge(dodge.width = 0.8))+
  geom_boxplot(outlier.colour = NA,notch=F,alpha=0.2)+
  scale_colour_manual(values = c('grey','darkorange'))+
  scale_fill_manual(values = c('grey','darkorange'))+
  annotate('text',x = 0.55,y = 2.1,label = 'B')

Fig3C <- ggplot(temp,
       aes(x = Tarsus,
           y = TROC,
           col=`Age (months)`)) +
  ylab('Loss in RTL') + 
  xlab('Tarsus length (mm)') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        legend.position = 'top') +
  geom_point(alpha = 0.2)+
  stat_smooth(method = 'lm')+
    scale_colour_manual(values = c(grey(0.2),'darkorange','darkgreen'))+
  scale_fill_manual(values = c('darkgrey','darkorange','darkgreen'))+
  annotate('text',x = 20,y = 2.1,label = 'C')

pdf('Figures/Figure 3.pdf',
    width = 8,
    height = 9,
    paper = 'a4r')
multiplot(Fig3A,Fig3B,Fig3C,layout = matrix(c(1,1,2,3),nrow = 2,byrow = T))
dev.off()

@@
